---
title: "Untitled"
author: "Daria Dodon"
date: "2024-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(here)
library(enrichR)
library(clusterProfiler)
library(readxl)   
library(tibble)   
library(janitor) 
library(org.Mm.eg.db) 
library(enrichplot)
library(RColorBrewer)
library(ggplot2)
```
```



```{r}
ensembl <- useEnsembl(biomart = "ensembl", 
                   dataset = "mmusculus_gene_ensembl", 
                   mirror = "useast")
```

```{r, include=FALSE}

data<- read.csv("DE_1m_oligodendrocytes.csv", row.names = 1)
```

```{r}
genes_entrez <- getBM(filters = "external_gene_name",
             attributes = c("entrezgene_id", "external_gene_name"),
             values = rownames(data),
             mart = ensembl)
```

```{r, include=FALSE}

data <- data %>% tibble::rownames_to_column("genes")
```

```{r, include=FALSE}

filtered_data <- data %>%
  filter(!grepl("^mt-", genes))
```

```{r, include=FALSE}

data_1 <- filtered_data %>% 
  mutate(ranking = (sign(avg_log2FC) * (-log10(p_val_adj)))) %>% 
  arrange(desc(ranking))
```

```{r, include=FALSE}


de_results <- subset(data_1, abs(pct.1 - pct.2) > 0.2)

```

```{r, include=FALSE}


de_results <- dplyr::left_join(de_results, genes_entrez, by = c("genes" = "external_gene_name"))
```

```{r, include=FALSE}

de_results %>% janitor::get_dupes("entrezgene_id")
```

```{r, include=FALSE}

de_results <- de_results %>%
  tidyr::drop_na() %>%
  dplyr::distinct(entrezgene_id, .keep_all = TRUE)
```


```{r}
top_genes <- de_results %>%
  arrange(desc(avg_log2FC)) %>%
  head(500)
```


```{r}
genes_list <- top_genes$ranking
names(genes_list) <- top_genes$entrezgene_id
genes_list <- sort(genes_list, decreasing = TRUE)
genes_list <- genes_list[is.finite(genes_list)]
```

```{r}

gseKEGG(genes_list, organism = "mmu", by = "DOSE", nPerm = 1000)
```


```{r}
gseWP_res_025 <- gseWP(genes_list, organism = "Mus musculus", by = "DOSE", nPerm = 1000, pvalueCutoff = 0.25)

```